\Large{Learning to Pivot with Adversarial Networks}

### 2016-11-02 tex/paper.tex
save

Many inference problems involve data generation processes that are not uniquely
specified or are uncertain in some way. In a scientific context, the presence of
several plausible data generation processes is often associated to the presence
of systematic uncertainties. Robust inference is possible if it is based on a
pivot -- a quantity whose distribution is invariant to the unknown value of the
(categorical or continuous) nuisance parameters that parametrizes this family of
generation processes. In this work,  we introduce a flexible training procedure
based on adversarial networks for enforcing the pivotal property on a predictive
model. We derive theoretical results proving that it
converges towards a predictive model that is both optimal and independent on the
nuisance parameters (if that models exists) or for which one can tune the
trade-off between power and robustness. Finally, we demonstrate the
effectiveness of this approach with a toy example and an example from particle
physics.

>>>>

Many inference problems involve data generation processes that are not uniquely
specified or are uncertain in some way. In a scientific context, the presence of
several plausible data generation processes is often associated to the presence
of systematic uncertainties. Robust inference is possible if it is based on a
pivot -- a quantity whose distribution is invariant to the unknown value of the
(categorical or continuous) nuisance parameters that parametrizes this family of
generation processes. In this work,  we introduce a flexible training procedure
based on adversarial networks for enforcing the pivotal property on a predictive
model. We derive theoretical results showing that the proposed algorithm tends towards a minimax
solution corresponding to a predictive model that is both optimal and
independent on the nuisance parameters (if that models exists) or for which one
can tune the trade-off between power and robustness. Finally, we demonstrate the
effectiveness of this approach with a toy example and an example from particle
physics.
